>

### 2019-04-09 paper/spinifex-paper.tex
polish and proofreading checks.

A tour is a multivariate data analysis technique in which a sequence of
linear (orthogonal) projections are viewed as an animation while the
orientation of the projection basis is rotated across time. Each frame
of the sequence corresponds to a small change in the projection for a
smooth transition that perseveres continuity. This is a crucial tool
extending the dimensionality of visualization which is a large boon as
datasets and models exist in increasing cardinality. The package
\textbf{tourr} (Wickham et al.
\protect\hyperlink{ref-wickham_tourr_2011}{2011}) builds a platform for
generating tour paths and applying to a wide number of display
geometrics in \textbf{base} graphics.

While there are numerous methods that generate tour paths, this research
focuses on the manual tour. The manual tour was described in Cook and
Buja (\protect\hyperlink{ref-cook_manual_1997}{1997}) and allows a user
to control the projection coefficients of a selected variable in a 2D
projection. The manipulation of these coefficients allows the analyst to
explore how sensitive the projections structure is to these changes.
This makes manual tours particularly useful once a feature of interest
has been identified, for example, with the use of a guided tour (Hurley
and Buja \protect\hyperlink{ref-hurley_analyzing_1990}{1990}). Rather
than exploring the large phase space of the data, a guided tour selects
a very specific path via projection pursuit(Diaconis and Freedman
\protect\hyperlink{ref-diaconis_asymptotics_1984}{1984}), the
optimization of an index function on the projection via a hill climbing
algorithm. This allows guided tours to rapidly identify interesting
projection features given the relatively large parameter-space. After a
projection of interest is identified an analyst can then use the ``finer
brush'' of the manual tour, by controlling the contributions of
individual variables.

Spinifex utilizes two new animation packages, \textbf{plotly} (Sievert
\protect\hyperlink{ref-sievert_plotly_2018}{2018}) and
\textbf{gganimate} (Pedersen and Robinson
\protect\hyperlink{ref-pedersen_gganimate:_2019}{2019}), to display
tours, manual or other saved tours. From a given projection, the user
can choose which variable to control, and the animation sequence is
generated to remove the variable from the projection, and then extend
its contribution to be the sole variable in one direction. This allows
the viewer to assess the change in structure induced in the projection
by the variable's contribution.

@@

{\centering \includegraphics[width=0.6\linewidth]{spinifex_paper_files/figure-latex/step0-1}

@@

Call \texttt{view\_basis()} on a basis to produce a \textbf{ggplot2}
graphic similar to \textbackslash ref\{fig:step0). Projection space is
always available for display via the matrix multiplication
\(\textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]} ~=~ \textbf{P}_{[n,~d]}\).

@@

# Also see vignette: 
vignette("spinifex") # vignette ‘spinifex’ not found

@@

{\centering \includegraphics[width=0.6\linewidth]{spinifex_paper_files/figure-latex/step2-1}

@@

Animations can be produced using the function
\texttt{play\_manual\_tour()}. This function defaults to an HTML5 widget
produced from \textbf{plotly}. The \texttt{render\_type} argument can be
changed to \texttt{render\_gganimate} for exporting to .gif or .mp4
files.

@@

The \textbf{tourr} package utilizes \textbf{base} graphics for the
display of tours. \textbf{spinifex} allows tours to be used in rendered
in \textbf{plotly} as an HTML5 object or \textbf{gganimate} as .gif or
.mp4 objects. Both of which build off \textbf{ggplot2} objects in
internal functions. Sharing of animations is not trivial especially in
print and static file formats such as .pdf. Even with the use of
computers and dynamic file formats capturing the correct resolution,
aspect, and display is challenging and many formats quickly bloat file
sizes. Keep in mind hosting options and exporting functions from
\textbf{plotly}, \textbf{gganimate} and \textbf{tourr}.

@@

The \textbf{tourr} package provides many other geometric displays with
the \texttt{tourr::display\_*()} family. These geometric options could
be integrated into the \textbf{ggplot2} framework for display on
\textbf{plotly} and \textbf{gganimate}. Additionally, the
\textbf{animation} package Xie et al.
(\protect\hyperlink{ref-xie_animation:_2018}{2018}) could be implemented
for another graphics framework. However, \textbf{animation} builds from
\textbf{base} graphs while \textbf{spinifex} utilizes \textbf{ggplot2}
graphics.

@@

Having a script only interaction with tours causes a significant barrier
to entry. To a lesser extent, \textbf{plotly} offers some static
interactions with the contained object, such as tooltips, brushing, and
linking without communicating back to the R console. The development of
a dynamic graphical user interface, perhaps with the use of a
\textbf{shiny} (Chang et al.
\protect\hyperlink{ref-chang_shiny:_2018}{2018}) application, would
mitigate the barrier to entry, allow for more rapid analysis, and offer
an approachable demo tool. The user could easily switch between
variables to control, adjust interpolation step angle, or flag/save
specific frame basis sets.

@@

This article was created in R (R Core Team
\protect\hyperlink{ref-r_core_team_r:_2018}{2018}), using
\textbf{bookdown} (Xie \protect\hyperlink{ref-xie_bookdown:_2016}{2016})
and \textbf{rmarkdown} (Xie, Allaire, and Grolemund
\protect\hyperlink{ref-xie_r_2018}{2018}), with code generating the
examples inline. The source files for this article be found at
\href{https://github.com/nspyrison/confirmation/}{github.com/nspyrison/confirmation/}.
The source code for the \textbf{spinifex} package can be found at
\href{https://github.com/nspyrison/spinifex/}{github.com/nspyrison/spinifex/}.

@@

@@

\leavevmode\hypertarget{ref-xie_bookdown:_2016}{}%
Xie, Yihui. 2016. \emph{Bookdown: Authoring Books and Technical
Documents with R Markdown}. Boca Raton, Florida: Chapman; Hall/CRC.
\url{https://github.com/rstudio/bookdown}.

@@

\address{%
Nicholas Spyrison\\
Monash University\\
School of Information Technology\\ Monash University, Melbourne, VIC 3800, Australia\\
}
\href{mailto:Nicholas.Spyrison@monash.edu}{\nolinkurl{Nicholas.Spyrison@monash.edu}}

\address{%
Dianne Cook\\
Monash University\\
School of Econometrics and Business Statistics\\ Monash University, Melbourne, VIC 3800, Australia\\
}
\href{mailto:DiCook@monash.edu}{\nolinkurl{DiCook@monash.edu}}

>>>>

A tour is a multivariate data analysis technique in which a sequence of
linear (orthogonal) projections are viewed as an animation while the
orientation of the projection basis is rotated across time. Each frame
of the sequence corresponds to a small change in the projection for a
smooth transition that perseveres continuity. This is a crucial tool
extending the dimensionality of visualization which is a large boon as
datasets and models exist in increasing cardinality. The package
\CRANpkg{tourr} (Wickham et al.
\protect\hyperlink{ref-wickham_tourr_2011}{2011}) builds a platform for
generating tour paths and applying to a wide number of display
geometrics in \textbf{base} graphics.

While there are numerous methods that generate tour paths, this research
focuses on the manual tour. The manual tour was described in Cook and
Buja (\protect\hyperlink{ref-cook_manual_1997}{1997}) and allows a user
to control the projection coefficients of a selected variable in a 2D
projection. The manipulation of these coefficients allows the analyst to
explore how sensitive the projections structure is to these changes.
This makes manual tours particularly useful once a feature of interest
has been identified, for example, with the use of a guided tour (Hurley
and Buja \protect\hyperlink{ref-hurley_analyzing_1990}{1990}). Rather
than exploring the large phase space of the data, a guided tour selects
a very specific path via projection pursuit (Diaconis and Freedman
\protect\hyperlink{ref-diaconis_asymptotics_1984}{1984}), the
optimization of an index function on the projection via a hill climbing
algorithm. This allows guided tours to rapidly identify interesting
projection features given the relatively large parameter-space. After a
projection of interest is identified an analyst can then use the ``finer
brush'' of the manual tour, by controlling the contributions of
individual variables.

\CRANpkg{spinifex} utilizes two new animation packages, \CRANpkg{plotly}
(Sievert \protect\hyperlink{ref-sievert_plotly_2018}{2018}) and
\CRANpkg{gganimate} (Pedersen and Robinson
\protect\hyperlink{ref-pedersen_gganimate:_2019}{2019}), to display
tours, manual or other saved tours. From a given projection, the user
can choose which variable to control, and the animation sequence is
generated to remove the variable from the projection, and then extend
its contribution to be the sole variable in one direction. This allows
the viewer to assess the change in structure induced in the projection
by the variable's contribution.

@@

{\centering \includegraphics[width=0.7\linewidth]{spinifex_paper_files/figure-latex/step0-1}

@@

Call \texttt{view\_basis()} on a basis to produce a \CRANpkg{ggplot2}
(Wickham \protect\hyperlink{ref-wickham_ggplot2:_2016}{2016}) graphic
similar to \textbackslash ref\{fig:step0). Projection space is always
available for display via the matrix multiplication
\(\textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]} ~=~ \textbf{P}_{[n,~d]}\).

@@

# Also see the vignette: 
vignette("spinifex")

@@

{\centering \includegraphics[width=0.7\linewidth]{spinifex_paper_files/figure-latex/step2-1}

@@

Animations can be produced using the function
\texttt{play\_manual\_tour()}. This function defaults to an HTML5 widget
produced from \textbf{plotly}. The \texttt{render\_type} argument can be
changed to \texttt{render\_gganimate} for exporting to GIF or MP4 files.

@@

The \textbf{tourr} package utilizes \textbf{base} graphics for the
display of tours. \textbf{spinifex} allows tours to be used in rendered
in \textbf{plotly} as an HTML5 object or \textbf{gganimate} as GIF or
MP4 files. Both of which build off \textbf{ggplot2} objects in internal
functions. Sharing of animations is not trivial especially in print and
static file formats such as PDF. Even with the use of computers and
dynamic file formats capturing the correct resolution, aspect, and
display is challenging and many formats quickly bloat file sizes. Keep
in mind hosting options and exporting functions from \textbf{plotly},
\textbf{gganimate} and \textbf{tourr}.

@@

The \textbf{tourr} package provides many other geometric displays with
the \texttt{tourr::display\_*()} family. These geometric options could
be integrated into the \textbf{ggplot2} framework for display on
\textbf{plotly} and \textbf{gganimate}. Additionally, the
\CRANpkg{animation} package Xie et al.
(\protect\hyperlink{ref-xie_animation:_2018}{2018}) could be implemented
for another graphics framework. However, \textbf{animation} builds from
\textbf{base} graphs while \textbf{spinifex} utilizes \textbf{ggplot2}
graphics.

@@

Having a script only interaction with tours causes a significant barrier
to entry. To a lesser extent, \textbf{plotly} offers some static
interactions with the contained object, such as tooltips, brushing, and
linking without communicating back to the R console. The development of
a dynamic graphical user interface, perhaps with the use of a
\CRANpkg{shiny} (Chang et al.
\protect\hyperlink{ref-chang_shiny:_2018}{2018}) application, would
mitigate the barrier to entry, allow for more rapid analysis, and offer
an approachable demo tool. The user could easily switch between
variables to control, adjust interpolation step angle, or flag/save
specific frame basis sets.

@@

This article was created in R (R Core Team
\protect\hyperlink{ref-r_core_team_r:_2018}{2018}), using
\CRANpkg{knitr} (Xie \protect\hyperlink{ref-stodden_knitr:_2014}{2014})
and \CRANpkg{rmarkdown} (Xie, Allaire, and Grolemund
\protect\hyperlink{ref-xie_r_2018}{2018}), with code generating the
examples inline. The source files for this article be found at
\href{https://github.com/nspyrison/spinifex_paper/}{github.com/nspyrison/spinifex\_paper/}.
The source code for the \textbf{spinifex} package can be found at
\href{https://github.com/nspyrison/spinifex/}{github.com/nspyrison/spinifex/}.

@@

\leavevmode\hypertarget{ref-wickham_ggplot2:_2016}{}%
Wickham, Hadley. 2016. \emph{Ggplot2: Elegant Graphics for Data
Analysis}. Springer-Verlag New York. \url{http://ggplot2.org}.

@@

\leavevmode\hypertarget{ref-stodden_knitr:_2014}{}%
Xie, Yihui. 2014. ``Knitr: A Comprehensive Tool for Reproducible
Research in R.'' In \emph{Implementing Reproducible Computational
Research}, edited by Victoria Stodden, Friedrich Leisch, and Roger D.
Peng. Chapman; Hall/CRC.
\url{http://www.crcpress.com/product/isbn/9781466561595}.

@@

\address{%
Nicholas Spyrison\\
Monash University\\
Faculty of Information Technology\\
}
\href{mailto:Nicholas.Spyrison@monash.edu}{\nolinkurl{Nicholas.Spyrison@monash.edu}}

\address{%
Dianne Cook\\
Monash University\\
Department of Econometrics and Business Statistics\\
}
\href{mailto:DiCook@monash.edu}{\nolinkurl{DiCook@monash.edu}}

### 2019-04-09 paper/spinifex_paper.Rmd
polish and proofreading checks.

---
title: > 
  spinifex: manual control of dynamic linear projections of high-dimensional data
author:
  - name: Nicholas Spyrison
    affiliation: Monash University 
    address:
    - School of Information Technology
    - Monash University, Melbourne, VIC 3800, Australia
    email:  Nicholas.Spyrison@monash.edu
  - name: Dianne Cook
    affiliation: Monash University 
    address:
    - School of Econometrics and Business Statistics
    - Monash University, Melbourne, VIC 3800, Australia
    email:  DiCook@monash.edu
abstract: >
  The set of dynamic linear projections of multivariate data collectively known as **tours** provides an important tool for extending the dimensionality of visuals. The R package **tourr** offers a variety of path generators and geometric displays for conducting tours. This paper discusses an extension package, **spinifex**, that adds support for the path generation of manual tours and extends the display of tours to use with the animation packages, **plotly** and **gganimate**. Manual tours are used to explore the sensitivity of structure as the contributions of a manipulation variable are changed.

@@

<!-- Tours -->
A tour is a multivariate data analysis technique in which a sequence of linear (orthogonal) projections are viewed as an animation while the orientation of the projection basis is rotated across time. Each frame of the sequence corresponds to a small change in the projection for a smooth transition that perseveres continuity. This is a crucial tool extending the dimensionality of visualization which is a large boon as datasets and models exist in increasing cardinality. The package **tourr** [@wickham_tourr_2011] builds a platform for generating tour paths and applying to a wide number of display geometrics in **base** graphics.

<!-- Manual tour -->
While there are numerous methods that generate tour paths, this research focuses on the manual tour. The manual tour was described in @cook_manual_1997 and allows a user to control the projection coefficients of a selected variable in a 2D projection. The manipulation of these coefficients allows the analyst to explore how sensitive the projections structure is to these changes. This makes manual tours particularly useful once a feature of interest has been identified, for example, with the use of a guided tour [@hurley_analyzing_1990]. Rather than exploring the large phase space of the data, a guided tour selects a very specific path via projection pursuit[@diaconis_asymptotics_1984], the optimization of an index function on the projection via a hill climbing algorithm. This allows guided tours to rapidly identify interesting projection features given the relatively large parameter-space. After a projection of interest is identified an analyst can then use the "finer brush" of the manual tour, by controlling the contributions of individual variables.

<!-- Display -->
Spinifex utilizes two new animation packages, **plotly** [@sievert_plotly_2018] and **gganimate** [@pedersen_gganimate:_2019], to display tours, manual or other saved tours. From a given projection, the user can choose which variable to control, and the animation sequence is generated to remove the variable from the projection, and then extend its contribution to be the sole variable in one direction. This allows the viewer to assess the change in structure induced in the projection by the variable's contribution.

@@

```{r step0, echo=F, results='hide', out.width='60%',                             fig.cap = "Basis reference axes (left) and projected data (right) of standardized flea data. Data points color and shape are mapped to beetle species. Basis identified by a holes-index guided tour. The variables aede2 and tars1 contribute mostly orthogonal to the other variables."}
## Flea holes tour
f_dat  <- tourr::rescale(flea[,1:6])
f_cat  <- factor(flea$species)
f_path <- save_history(f_dat, guided_tour(holes()))
f_bas  <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2)
f_mvar <- 5
f_msp  <- create_manip_space(basis = f_bas, manip_var = f_mvar)
f_proj <- data.frame(tourr::rescale(f_dat %*% f_msp[, 1:2]))

@@

Call `view_basis()` on a basis to produce a **ggplot2** graphic similar to \ref{fig:step0). Projection space is always available for display via the matrix multiplication $\textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]} ~=~ \textbf{P}_{[n,~d]}$.

@@

```{r step2, echo=F, out.width='60%',                                           fig.cap = "Manipulation space for controlling the contribution of aede2 of standardized flea data. Basis selected by a holes-index optimized guided tour lies on the horizontal projection plane, shown in blue. The manip var axis, in red, extends into the vertical manipulation space, allowing the coefficients of the manip var to be changed by rotation around the origin."}

@@

```{r step3, echo=F, warning=F, fig.height=7.2, fig.width=6, out.height='7.2in', out.width='6in',                                                                fig.cap = "Radial manual tour manipulating aede2 of standardized flea data.  The contributions increase from its initial contribution to a full contribution to the projection before decreasing to zero and then returning to its initial value. The change in the projected data shows that aede2 is important for distinguishing the purple species. An animated version can be viewed at  https://nspyrison.netlify.com/thesis/flea\\_manualtour\\_mvar5/."}
f_angle <- .26
f_mtour <- manual_tour(f_bas, manip_var = f_mvar, angle = f_angle)
if(dim(f_mtour)[3] != 15) message(
  paste0("!!!!! step3 n_slides != 15 !!!!! n_slides = ", dim(f_mtour)[3]))

@@

Animations can be produced using the function `play_manual_tour()`. This function defaults to an HTML5 widget produced from **plotly**. The `render_type` argument can be changed to `render_gganimate` for exporting to .gif or .mp4 files.

@@

The **tourr** package utilizes **base** graphics for the display of tours. **spinifex** allows tours to be used in rendered in **plotly** as an HTML5 object or **gganimate** as .gif or .mp4 objects. Both of which build off **ggplot2** objects in internal functions. Sharing of animations is not trivial especially in print and static file formats such as .pdf. Even with the use of computers and dynamic file formats capturing the correct resolution, aspect, and display is challenging and many formats quickly bloat file sizes. Keep in mind hosting options and exporting functions from **plotly**, **gganimate** and **tourr**.

@@

<!-- geometric and graphics display -->
The **tourr** package provides many other geometric displays with the `tourr::display_*()` family. These geometric options could be integrated into the **ggplot2** framework for display on **plotly** and **gganimate**. Additionally, the **animation** package @xie_animation:_2018 could be implemented for another graphics framework. However, **animation** builds from **base** graphs while **spinifex** utilizes **ggplot2** graphics.

@@

<!-- dynamic interaction/gui -->
Having a script only interaction with tours causes a significant barrier to entry. To a lesser extent, **plotly** offers some static interactions with the contained object, such as tooltips, brushing, and linking without communicating back to the R console. The development of a dynamic graphical user interface, perhaps with the use of a **shiny** [@chang_shiny:_2018] application, would mitigate the barrier to entry, allow for more rapid analysis, and offer an approachable demo tool. The user could easily switch between variables to control, adjust interpolation step angle, or flag/save specific frame basis sets.

@@

This article was created in R [@r_core_team_r:_2018], using **bookdown** [@xie_bookdown:_2016] and **rmarkdown** [@xie_r_2018], with code generating the examples inline. The source files for this article be found at [github.com/nspyrison/confirmation/](https://github.com/nspyrison/confirmation/). The source code for the **spinifex** package can be found at [github.com/nspyrison/spinifex/](https://github.com/nspyrison/spinifex/).

>>>>

---
title: > 
  spinifex: manual control of dynamic linear projections of high-dimensional data
author:
  - name: Nicholas Spyrison
    affiliation: Monash University 
    address:
    - Faculty of Information Technology
    email:  Nicholas.Spyrison@monash.edu
  - name: Dianne Cook
    affiliation: Monash University 
    address:
    - Department of Econometrics and Business Statistics
    email:  DiCook@monash.edu
abstract: >
  The set of dynamic linear projections of multivariate data collectively known as **tours** provides an important tool for extending the dimensionality of visuals. The R package **tourr** offers a variety of path generators and geometric displays for conducting tours. This paper discusses an extension package, **spinifex**, that adds support for the path generation of manual tours and extends the display of tours to use with the animation packages, **plotly** and **gganimate**. Manual tours are used to explore the sensitivity of structure as the contributions of a manipulation variable are changed.

@@

<!-- Tours -->
A tour is a multivariate data analysis technique in which a sequence of linear (orthogonal) projections are viewed as an animation while the orientation of the projection basis is rotated across time. Each frame of the sequence corresponds to a small change in the projection for a smooth transition that perseveres continuity. This is a crucial tool extending the dimensionality of visualization which is a large boon as datasets and models exist in increasing cardinality. The package \CRANpkg{tourr} [@wickham_tourr_2011] builds a platform for generating tour paths and applying to a wide number of display geometrics in **base** graphics.

<!-- Manual tour -->
While there are numerous methods that generate tour paths, this research focuses on the manual tour. The manual tour was described in @cook_manual_1997 and allows a user to control the projection coefficients of a selected variable in a 2D projection. The manipulation of these coefficients allows the analyst to explore how sensitive the projections structure is to these changes. This makes manual tours particularly useful once a feature of interest has been identified, for example, with the use of a guided tour [@hurley_analyzing_1990]. Rather than exploring the large phase space of the data, a guided tour selects a very specific path via projection pursuit [@diaconis_asymptotics_1984], the optimization of an index function on the projection via a hill climbing algorithm. This allows guided tours to rapidly identify interesting projection features given the relatively large parameter-space. After a projection of interest is identified an analyst can then use the "finer brush" of the manual tour, by controlling the contributions of individual variables.

<!-- Display -->
\CRANpkg{spinifex} utilizes two new animation packages, \CRANpkg{plotly} [@sievert_plotly_2018] and \CRANpkg{gganimate} [@pedersen_gganimate:_2019], to display tours, manual or other saved tours. From a given projection, the user can choose which variable to control, and the animation sequence is generated to remove the variable from the projection, and then extend its contribution to be the sole variable in one direction. This allows the viewer to assess the change in structure induced in the projection by the variable's contribution.

@@

```{r step0, echo=F, results='hide', out.width='70%',                             fig.cap = "Basis reference axes (left) and projected data (right) of standardized flea data. Data points color and shape are mapped to beetle species. Basis identified by a holes-index guided tour. The variables aede2 and tars1 contribute mostly orthogonal to the other variables."}
## Flea holes tour
f_dat  <- tourr::rescale(flea[,1:6])
f_cat  <- factor(flea$species)
f_path <- save_history(f_dat, guided_tour(holes()))
f_bas  <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2)
f_mvar <- 5
f_msp  <- create_manip_space(basis = f_bas, manip_var = f_mvar)
f_proj <- data.frame(tourr::rescale(f_dat %*% f_msp[, 1:2]))

@@

Call `view_basis()` on a basis to produce a \CRANpkg{ggplot2} [@wickham_ggplot2:_2016] graphic similar to \ref{fig:step0). Projection space is always available for display via the matrix multiplication $\textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]} ~=~ \textbf{P}_{[n,~d]}$.

@@

```{r step2, echo=F, out.width='70%',                                           fig.cap = "Manipulation space for controlling the contribution of aede2 of standardized flea data. Basis selected by a holes-index optimized guided tour lies on the horizontal projection plane, shown in blue. The manip var axis, in red, extends into the vertical manipulation space, allowing the coefficients of the manip var to be changed by rotation around the origin."}

@@

```{r step3, echo=F, warning=F, fig.height=7.2, fig.width=6, out.height='7.2in', out.width='6in',                                                                fig.cap = "Radial manual tour manipulating aede2 of standardized flea data.  The contributions increase from its initial contribution to a full contribution to the projection before decreasing to zero and then returning to its initial value. The change in the projected data shows that aede2 is important for distinguishing the purple species. An animated version can be viewed at  https://nspyrison.netlify.com/thesis/flea\\_manualtour\\_mvar5/."}
f_angle <- .26
f_mtour <- manual_tour(f_bas, manip_var = f_mvar, angle = f_angle)
if(dim(f_mtour)[3] != 15) message(
  paste0("!!!!! step3 n_slides != 15 !!!!! Please knit again. n_slides = ", dim(f_mtour)[3]))

@@

Animations can be produced using the function `play_manual_tour()`. This function defaults to an HTML5 widget produced from **plotly**. The `render_type` argument can be changed to `render_gganimate` for exporting to GIF or MP4 files.

@@

The **tourr** package utilizes **base** graphics for the display of tours. **spinifex** allows tours to be used in rendered in **plotly** as an HTML5 object or **gganimate** as GIF or MP4 files. Both of which build off **ggplot2** objects in internal functions. Sharing of animations is not trivial especially in print and static file formats such as PDF. Even with the use of computers and dynamic file formats capturing the correct resolution, aspect, and display is challenging and many formats quickly bloat file sizes. Keep in mind hosting options and exporting functions from **plotly**, **gganimate** and **tourr**.

@@

<!-- geometric and graphics display -->
The **tourr** package provides many other geometric displays with the `tourr::display_*()` family. These geometric options could be integrated into the **ggplot2** framework for display on **plotly** and **gganimate**. Additionally, the \CRANpkg{animation} package @xie_animation:_2018 could be implemented for another graphics framework. However, **animation** builds from **base** graphs while **spinifex** utilizes **ggplot2** graphics.

@@

<!-- dynamic interaction/gui -->
Having a script only interaction with tours causes a significant barrier to entry. To a lesser extent, **plotly** offers some static interactions with the contained object, such as tooltips, brushing, and linking without communicating back to the R console. The development of a dynamic graphical user interface, perhaps with the use of a \CRANpkg{shiny} [@chang_shiny:_2018] application, would mitigate the barrier to entry, allow for more rapid analysis, and offer an approachable demo tool. The user could easily switch between variables to control, adjust interpolation step angle, or flag/save specific frame basis sets.

@@

This article was created in R [@r_core_team_r:_2018], using \CRANpkg{knitr} [@stodden_knitr:_2014] and \CRANpkg{rmarkdown} [@xie_r_2018], with code generating the examples inline. The source files for this article be found at [github.com/nspyrison/spinifex_paper/](https://github.com/nspyrison/spinifex_paper/). The source code for the **spinifex** package can be found at [github.com/nspyrison/spinifex/](https://github.com/nspyrison/spinifex/).
